---
title: "R Notebook"
output: html_notebook
---

A Portuguese banking institution is interested in unleashing a marketing campaign for the subscription of its term deposits. The marketing campaign will be focussed towards the existing customers of the banking institution. The bank does have historical data of its customers encompassing a wide range of attributes (Client information, previous cam-paign and other socio economic attributes). 

In the interest of spending money judiciously, the bank would like to know beforehand the possibility of a customer subscribing to its term deposit. With this knowledge, the bank can target a select pool of customers, thereby reducing its marketing campaign ex-penses and at the same time getting a higher conversion ratio (Number of customers sub-scribing to the term deposit over the number of customers contacted).

Armed with the historical data about the bank’s clients and the previous marketing cam-paigns, the objective is to develop a Classification (Logistic Regression) algorithm which would predict if a customer would or would not subscribe to a term deposit. Thus, this would be a binary (yes/no) classification problem with 20 predictor variables to start with.

Data can be found at https://archive.ics.uci.edu/ml/datasets/Bank+Marketing#


```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("/Users/anon/GitHubRepos/CourseFiles/Classification/")) 
```


```{r include = FALSE}
getwd()
```

```{r message=FALSE}

library(ggplot2)
library(caret)
library(dummies)
library(reshape2)
library(unbalanced)
library(ROCR)
library(rms)
```

```{r}

#Reading the file. Sep = ; was used as with , all variables were coming under a single column
bank_data <- read.csv("bank-additional-full.csv", header = TRUE, sep = ";", na.strings = c(""))

```

```{r}
#Removing duration as instructed. its value is known only after call is made
bank_data$duration <- NULL
```

```{r}
#finding missing values
sapply(bank_data, function(x) sum(is.na(x)))
```

```{r}

#creating boxplot for the numeric variables for exploration
melt_bank_data <- melt(bank_data[,sapply(bank_data,is.numeric)])
ggplot(data=melt_bank_data,aes(x=variable,y=value))+geom_boxplot()+facet_wrap(~variable,scales="free")+theme_bw()

```

```{r}

#Binning Age numeric variable into age_binned which is a factor variable with 6 levels
bank_data$age_binned <- cut(bank_data$age, c(16,33,39,48,98))

```

```{r}

#Converting previous and pdays values to contacted and non contacted as there is a 
#heavy skew towards zero and 999 value
bank_data$previous_cat <- ifelse(bank_data$previous==0,'Not Contacted','Contacted')
bank_data$pdays_cat <- ifelse(bank_data$pdays==999,'Not Contacted','Contacted')
bank_data$previous_cat <- as.factor(bank_data$previous_cat)
bank_data$pdays_cat <- as.factor(bank_data$pdays_cat)

```

```{r}

#Creating the (log) transformed campaign variable
bank_data$log_campaign <- log(1+bank_data$campaign)

```

```{r include=FALSE}



```

```{r include=FALSE}

```

```{r}

#Grouping three basic levels in education to one basic
basic <- c("basic.4y","basic.6y","basic.9y")
bank_data$education <- as.character(bank_data$education)
bank_data$education[bank_data$education %in% basic]<- 'basic'
bank_data$education<-as.factor(bank_data$education)
```

```{r}

#Converting yes to 1 and no to 0 in target variable y
bank_data$y <- ifelse(bank_data$y == "yes",1,0)

#dropping columns which are not required. variables dropped because of their transformation
columns_to_be_dropped <- c("age","campaign","pdays","previous")
bank_data_processed <- bank_data[,!names(bank_data) %in% columns_to_be_dropped]
```

```{r}

#ggplot for numeric predictors and target 
data_long = melt(bank_data_processed[, sapply(bank_data_processed, is.numeric)], id='y')

ggplot(data_long, aes(x = value, group=y, color=factor(y)) )+ 
  geom_density()+ facet_wrap(~variable, scales="free")

```

```{r}

#Creating dummy variables for levels of all factor variables
bank_data_dummy <- dummy.data.frame(bank_data_processed,names = c("job","education","marital","housing","loan","default","contact","month","day_of_week","poutcome","previous_cat","pdays_cat","age_binned"))

#Creating training and test sets
set.seed(100)
sample_list <- sample.int(nrow(bank_data_dummy), size = floor(0.7*nrow(bank_data_dummy)), replace = FALSE)
train_data_unbalanced <- bank_data_dummy[sample_list,]
test_data <- bank_data_dummy[-sample_list,]
```

```{r}

#Data balancing using ubSmote
set.seed(98765)
balanced <- ubSMOTE(train_data_unbalanced[,-which(names(train_data_unbalanced)=='y')],as.factor(train_data_unbalanced$y),perc.over = 200,k=5,perc.under = 200)
train_data_balanced <- cbind(balanced$X,Class=balanced$Y)

```

```{r}

#implementing the model
bank_model <- glm(Class~., data = train_data_balanced, family = binomial)
summary(bank_model)
```


```{r}

#Model iteration
bank_model_1 <- glm(Class~jobadmin.+jobretired+educationbasic+educationhigh.school+educationuniversity.degree+defaultno+housingno+contactcellular+monthapr+monthaug+monthdec+monthjul+monthjun+monthmar+monthmay+monthnov+monthoct+day_of_weekthu+poutcomefailure+emp.var.rate+cons.conf.idx+cons.price.idx+euribor3m+pdays_catContacted+log_campaign,data=train_data_balanced,family=binomial)
summary(bank_model_1)
```

```{r}

#Predicting on the training set and finding accuracy using confusion matrix
pred_train <- predict(bank_model, train_data_balanced, type = "response")
pred_train[pred_train<=0.5]=0
pred_train[pred_train>0.5]=1
confusionMatrix(pred_train,train_data_balanced$Class,positive = '1')

```

```{r}

pred_train_1 <- predict(bank_model_1, train_data_balanced, type = "response")
pred_train_1[pred_train_1<=0.5]=0
pred_train_1[pred_train_1>0.5]=1
confusionMatrix(pred_train_1,train_data_balanced$Class,positive = '1')
```

```{r}

#Predicting on the test set and finding accuracy using confusion matrix
pred_test <- predict(bank_model, test_data, type = "response")
pred_test[pred_test<=0.5]=0
pred_test[pred_test>0.5]=1
CM <- confusionMatrix(pred_test,test_data$y,positive = '1')
CM
```

```{r}
pred_test_1 <- predict(bank_model_1, test_data, type = "response")
pred_test_1[pred_test_1<=0.2]=0
pred_test_1[pred_test_1>0.2]=1
CM <- confusionMatrix(pred_test_1,test_data$y,positive = '1')
CM
```

```{r}

#Finding area under curve for test set prediction and plotting the curve too
pred_auc <- prediction(pred_test_1,test_data$y)
perf <- performance(pred_auc, measure = "tpr", x.measure = "fpr")
plot(perf,main="AUC on test prediction")
abline(a=0,b=1,lwd=2,lty=2,col="red")

auc <- performance(pred_auc, measure = "auc")
auc_value <- auc@y.values
round(as.numeric(auc_value), digits = 4)
```

```{r}

#Trend of specificity, sensitivty and accuracy with threshold
value_sensitivity <- vector()
value_specificity <- vector()
value_accuracy <- vector()


for (i in seq(0.1,0.5,0.05)) {
  pred_test_1 <- predict(bank_model_1, test_data, type = "response")
  pred_test_1[pred_test_1<=i]=0
  pred_test_1[pred_test_1>i]=1
  CM <- confusionMatrix(pred_test_1,test_data$y,positive = '1')
  value_sensitivity<-c(value_sensitivity,CM[[4]][1])
  value_specificity<-c(value_specificity,CM[[4]][2])
  value_accuracy<- c(value_accuracy,(CM[[3]][1]))
}
plot(seq(0.1,0.5,0.05),value_accuracy,col="red",main="Change in Accuracy over different threshold values",xlab="Threshold value",ylab="Accuracy")
lines(seq(0.1,0.5,0.05),value_accuracy)

plot(seq(0.1,0.5,0.05),value_sensitivity,col="red",main="Change in Sensitivity over different threshold values",xlab="Threshold value",ylab="Sensitivity")
lines(seq(0.1,0.5,0.05),value_sensitivity)

plot(seq(0.1,0.5,0.05),value_specificity,col="red",main="Change in Specificity over different threshold values",xlab="Threshold value",ylab="Specificity")
lines(seq(0.1,0.5,0.05),value_specificity)

```



